### 2025-03-16 manuscripts/theory/theory.qmd

---
title: "Behavioral Insight Design: A Psychological Framework for Designing Data-Driven Decision Support Systems"
subtitle: Behavioral Insight Design
author:
  - name: Jeremy R. Winget
    email: contact@jrwinget.com
    affiliations: 
        - id: some-tech
          name: Some Institute of Technology
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Milena Eickhoff
    email: m.eickhoff@skimgroup.com
    affiliations:
        - id: another-u
          name: Another University
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    note: |
      Another author footnote, this is a very long footnote and it should be a really long footnote. But this footnote is not yet sufficiently long enough to make two lines of footnote text.
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      # model: 3p # Don't set a model with preprint
      cite-style: authoryear
---

# Abstract 
- Define the Behavioral Insight Design (BID) framework. 
- Highlight theoretical foundations (cognitive psychology, behavioral economics, dual-process theories). 
- Emphasize its operationalization of psychological theory into practical design principles. 
- State its interdisciplinary value between psychology and UX/HCI.

# Introduction 
- Motivation: increased cognitive demands in data-rich environments. 
- Identify the gap: lack of structured psychological frameworks in UX/HCI. 
- Introduce BID as a synthesis of established psychological theories.

# Theoretical Background 
Summarize foundational theories relevant to BID: 
- Cognitive Load Theory (Sweller, 1988) 
- Dual-Process Theory (Kahneman & Tversky) 
- Cognitive Biases (Anchoring, Framing, Confirmation Bias) 
- Information Processing and Visual Perception (Gestalt principles, Tufte) 
- Peak–End Rule & Aesthetic–Usability Effect (Dion et al., 1972; Norman, 2002)

# The Behavioral Insight Design (BID) Framework 
Outline BID’s 5-stage model explicitly mapping psychological theories to each: 
- Stage 1: Notice (Cognitive Load, Hick’s Law, Visual Hierarchy) 
- Stage 2: Interpret (Data Storytelling, Processing Fluency, Emotional Framing) 
- Stage 3: Structure (Gestalt Principles, Dual-Processing, Default Effects) 
- Stage 4: Anticipate (Anchoring, Framing, Confirmation Bias, Risk Perception) 
- Stage 5: Validate & Empower (Peak–End Rule, Beautiful-is-Good, Cooperation & Coordination)

# Discussion: Theoretical Implications & Innovations 
- How BID operationalizes psychological concepts into UX steps. 
- Novelty in explicitly anticipating and mitigating biases within design phases. 
- Potential as a tool for cognitive/behavioral psychology experimentation.

# Future Directions 
- Propose empirical studies testing BID-derived predictions. 
- Suggest deeper investigations into psychological mechanisms influencing dashboard efficacy (e.g., cognitive load, memory retention, decision accuracy).

# Conclusion
- Restate BID’s theoretical contribution as a coherent synthesis of psychology into design methodology.