### 2025-04-04 conference-materials/2025_shiny-conf/slides.qmd
Updates {bidux} examples and refines results slide
- Enhances {bidux} code examples to improve clarity and output presentation
- Refines results slide to emphasize real-world impact while balancing the text on slide
- Updates package dependencies in renv.lock to include necessary packages for {bidux} functionality.

---
title: "Death By Dropdown?"
subtitle: "A Developer's Guide To Building Dashboards That Won't Fry Your Client's Brain"
author: "Milena Eickhoff & Jeremy Winget, PhD"
format: 
  revealjs:
    theme: simple
    slide-number: false
    footer: "ShinyConf 2025 | Death By Dropdown?"
    incremental: true
    transition: slide
    background-transition: fade
    highlight-style: github
    code-block-background: true
---

## The Problem {.center background-color="#4682b4"}

:::: {.columns}
::: {.column .nonincremental width="60%"}
<h3>Ever felt like this?</h3>

- Endless dropdown menus
- Overwhelming complexity
- Users disengaged
- Delayed decision-making
- "Dashboard rot"
:::

::: {.column width="40%"}
![](https://media0.giphy.com/media/v1.Y2lkPTc5MGI3NjExbzVhM3d3MjJvYTNyc3puM2E0cmduZjFweXhrczNwcXI0b3FhN203aCZlcD12MV9pbnRlcm5hbF9naWZfYnlfaWQmY3Q9Zw/l44Q5OXJ6qaNr838Q/giphy.gif)
:::
::::

::: {.notes}
**Opening Hook**: Begin with energy! Make eye contact across the room.

- "Raise your hand if you've ever encountered a dashboard that looked like this..." (gesture to image)
- "Dashboard rot" = dashboards built but never used - costing real development time and money
- Ask: "How many of you have abandoned using a dashboard because it was too complex?" (engage audience)
:::

## It Doesn't Have To Be This Way {.center}

<br>

::: {.r-fit-text}
By shifting the focus from <span style="color:#d55e00">**cramming data**</span> to <span style="color:#009e73">**crafting stories**</span>
:::

:::: {.columns .fragment}
::: {.column width="60%"}
<br>

SoundPulse's market research team was drowning in data, unable to extract meaningful insights from their complex dashboards.
:::

::: {.column width="40%"}
<br>

![](img/soundpulse-logo.png){width="100%"}
:::
::::

::: {.notes}
**Transition to Solution**: Shift to a hopeful, inspiring tone here.

- Key message: "The answer isn't more data - it's better storytelling with data"
- "SoundPulse, a music streaming analytics team, faced this exact problem"
  * They're a composite client
- Their market research team needed to present competitor analysis to executives
  * Typically involves turning tons of complex analyses into non-technical insights
- "We're going to show you exactly how psychological principles transformed their dashboard - with approaches and code you can use immediately"
:::

## BID In Action: Dashboard Transformation {.center background-color="#f2f2f2"}

::: {.absolute top="80%" left="12%" width="80%"}
![](img/before-bid.png)
:::

::: {.notes}
**Show Impact**: Build frustration recognition with the "before" dashboard.

- Even though SoundPulse is a composite client...this is representative of real dashboards we've seen across the market research industry
- Point out key issues:
  * "Filter overload before seeing ANY insights"
  * "Multiple tabs with no clear starting point"
  * "Complete cognitive overload"
- "This is why dashboards get abandoned - nobody wants this experience"
:::

## Behavior Insight Design (BID) Framework {background-color="#e6f0ff"}

::: {.absolute left="7%" width="90%"}
![](img/bid-framework.png)
:::

:::: {.columns .absolute top="65%" left="4%" .r-fit-text}

::: {.column width="38%"}
1. <span style="color:#1c6d7d">**Notice**</span> the Problem
:::

::: {.column width="46%"}
2. <span style="color:#35a4ae">**Interpret**</span> the User's Need
3. <span style="color:#ff9000">**Structure**</span> the Dashboard
4. <span style="color:#ffca90">**Anticipate**</span> User Behavior
:::

::: {.column width="16%"}
5. <span style="color:#98d3ca">**Validate**</span> & Empower the User
:::

::::

::: {.notes}
**Framework Introduction**: Slow down here - this is the core methodology.

- "The 5 stages form an easy workflow: Notice → Interpret → Structure → Anticipate → Validate"
  * Click through each
- "Each stage translates psychological principles into practical developer actions"
- "This isn't just theory - it's a step-by-step process you can implement today"
- "These stages are iterative - you'll often revisit earlier stages as you discover new insights"
  * Any order
:::

## Where BID Fits In {background-color="#f2f2f2"}

::: {.d-flex .align-items-center}
![](img/newton-portrait.jpg){width="15%" style="float:left; margin-right:15px;"}

> "If I have seen further, it is by standing on the shoulders of giants."
>
> — Isaac Newton
:::

:::: {.columns}

::: {.column}
- **UX Design**: Beyond heuristic lists to guided process
- **Data Storytelling**: From data dumps to meaningful narratives
:::

::: {.column}
- **Psychological Science**: Evidence-based cognitive principles
- **Visual Communication**: Optimized for information perception
:::

::::

::: {.callout-note .mt-3}
**Key Differentiator**: Process-oriented framework vs. typical checklist approach
:::

::: {.notes}
**Context Setting**: Position the framework within the broader fields.

- "BID integrates existing best practices from multiple disciplines"..."We're truly standing on the shoulders of giants here"
  * Click through example fields...UX, data storytelling, psych science, data/visual communication
- Key differentiator: "But what makes BID unique is its process-oriented framework vs. a typical checklist approach"
- "The goal is helping users gain clearer insights while giving developers a tool to use in their workflows"
- "Even though there's a lot of theory, you don't need a psychology degree to apply these principles effectively"
- "To demonstrate...let's take a closer look at how we connected these stages to SoundPulse's dashboard."
:::

## SoundPulse Case Study: BID Stages 1-3 {background-color="#f9f9f9"}

:::: {.columns .nonincremental .r-fit-text}

::: {.column width="32%"}
- <span style="color:#1c6d7d">**Stage 1: Notice**</span>
  * Complex filtering overwhelmed users (*Cognitive Load*)
  * 72% users reported confusion (*Hick's Law*)
:::

::: {.column width="32%"}
- <span style="color:#35a4ae">**Stage 2: Interpret**</span>
  * Intuitive visualizations (*Processing Fluency*)
  * Clearer key insights<br>(*Data Storytelling*)
:::

::: {.column width="32%"}
- <span style="color:#ff9000">**Stage 3: Structure**</span>
  * Group related elements (*Principle of Proximity*)
  * Prioritization of metrics (*Dual-Processing Theory*)
:::

::::

::: {.absolute bottom="15%" left="15%" width="70%" height="30%"}
![](img/bid-framework.png)
:::

::: {.notes}
**Case Study Introduction:** Keep energy high while introducing concepts.

- **Stage 1 (Notice):** The first thing was users were overwhelmed with options
  * This is what we call 'cognitive load' - when the brain gets overtaxed with information
  * We was a violation of Hick's Law, which simply means that more choices lead to longer decision times and confusion
- **Stage 2 (Interpret):** We then focused on something called 'Processing Fluency' - making the information easier to digest
  * For example, using color to create a visual hierarchy of information so the brain can process it more easily
- **Stage 3 (Structure):** "We then restructured the UI using the 'Principle of Proximity' - which is a fancy way of saying we started grouping related elements together. 
  * Now there are a ton of ways you can do this...
  * But as you'll see in our code example, we created groups of primary and secondary metrics to create a natural flow through the data
- Now we'll show you the actual code implementation of these principles
:::

## {reactable} Implementation: Visual Result {background-color="#f9f9f9"}

::: {.r-stack style="text-align:center"}
![](img/reactable-ex1.png){width="100%"}

![](img/reactable-ex2.png){.fragment width="100%"}

![](img/reactable-ex3.png){.fragment width="100%"}

![](img/reactable-ex4.png){.fragment width="100%"}

![](img/reactable-ex5.png){.fragment width="100%"}
:::

::: {.notes}
**Visual Impact**: Connect visuals to principles just mentioned.

- For each image as it appears:
  * First image: "Notice how grouping reduces cognitive load"
  * Second image: "The consistent legend simplifies comprehension"
  * Third image: "Color coding creates instant visual processing"
  * Fourth image: "Proximity principles group related metrics"
- "This is the power of intentional design - every element has a psychological purpose"
:::

## {reactable} Implementation: Basics {background-color="#f9f9f9"}

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: "4-5|6-18|19-23"

# Packages: reactable, dplyr, bslib
reactable(
  soundpulse_data,
  # Stage 1: Reduce cognitive load with grouped structure
  groupBy = "Song",
  # Stage 2: Enhance processing fluency with color coding
  columns = list(
    `Metric 1` = colDef(
      style = function(value) {
        # Color coding for instant comprehension
        case_when(
          value > 60 ~ list(background = "#1c6d7d", color = "white"),
          value > 40 ~ list(background = "#35a4ae", color = "white"),
          TRUE ~ list(background = "#ffca90", color = "black")
        )
      }
    )
  ),
  # Stage 1: Progressive disclosure reduces overload
  details = function(index) {
    song_details <- filter(details_data, song_id == soundpulse_data$Song[index])
    card(reactable(song_details, outlined = TRUE))
  }
)
```

::: {.notes}
**Basic Implementation**: Highlight the simplicity of applying BID.

- "This first example shows how just a few lines implement powerful BID principles"
- "groupBy reduces cognitive load while creating visual hierarchy"
- "Color coding isn't just aesthetic - it creates instant processing fluency"
- "Progressive disclosure means users only see details when needed"
- "These three techniques alone dramatically improve dashboard usability"
- "This simple implementation already reduces cognitive load and enhances processing - but we can do more"
:::

## {reactable} Implementation: Advanced {background-color="#f9f9f9"}

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: "4-14|15-49"

# Packages: reactable, dplyr, bslib
reactable(
  soundpulse_data,
  # Stage 3: Group related elements (Principle of Proximity)
  columnGroups = list(
    colGroup(
      name = "Core Metrics",
      columns = c("Metric 1", "Metric 2", "Metric 3", "Metric 4")
    ),
    colGroup(
      name = "Key Performance",
      columns = c("KPI 1", "KPI 2", "KPI 3")
    )
  ),
  # Stage 1: Dynamic legend reduces choice complexity (Hick's Law)
  defaultColDef = colDef(
    header = function(value) {
      layout_inline(
        value,
        span(
          style = css(
            display = "inline-block",
            margin_left = "5px",
            height = "10px", 
            width = "10px",
            border_radius = "50%",
            background_color = "#1c6d7d"
          )
        )
      )
    }
  ),
  # Combine multiple visual elements for age groups
  columns = list(
    Age = colDef(
      cell = function(value) {
        colors <- c(
          "18-24" = "#1c6d7d",
          "25-34" = "#35a4ae", 
          "35-44" = "#98d3ca",
          "45-54" = "#ffca90",
          "55-64" = "#1c6d7d",
          "65-74" = "#dd8500"
          )

        span(value, style = css(color = colors[value], font_weight = "bold"))
      }
    )
  )
)
```

::: {.notes}
**Advanced Features**: Highlight sophisticated techniques of package, but balance with BID principles.

- "columnGroups implements Principle of Proximity - grouping primary and secondary metrics together"
- "The dynamic legend reduces choice complexity - addressing Hick's Law"
- "Color coding by demographic creates instant pattern recognition"
- "Notice how each feature directly maps to a BID principle"
- Ask: "How many have used columnGroups in reactable before?" (engagement opportunity)
  * If not many...great, we taught something!
  * If a lot...could also mention (quickly) how to extend with legend/div
:::

## SoundPulse Case Study: BID Stages 4-5 {background-color="#f9f9f9"}

:::: {.columns .nonincremental .r-fit-text}

::: {.column width="48%"}
- <span style="color:#ffca90">**Stage 4: Anticipate**</span>
  * Executives comparing metrics without context (*Anchoring Effect*)
  * Teams interpreting data differently (*Confirmation Bias*)
  * Different perspectives needed for different teams (*Framing & Loss Aversion*)
:::

::: {.column width="48%"}
- <span style="color:#98d3ca">**Stage 5: Validate**</span>
  * Teams needed actionable summaries (*Peak-End Rule*)
  * No clear path to consensus<br>(*Cooperation & Coordination*)
:::
::::

::: {.absolute bottom="15%" left="15%" width="70%" height="30%"}
![](img/bid-framework.png)
:::

::: {.notes}
**Later stages:** Set up the psychological depth of later stages

- "For the later stages of the BID framework, we addressed more subtle psychological factors that affect how users interpret and act on data."
- **Stage 4 (Anticipate):** We noticed executives were making flawed comparisons without proper context.
  * This relates to the 'Anchoring Effect' - where people rely too heavily on the first piece of information they see
  * We also saw teams struggling to interpret the data in the same way, so we addressed this with something called a "framing toggle"...more on that in just a moment
- **Stage 5 (Validate):** Finally, we applied the 'Peak-End Rule' - the idea that people judge experiences mainly by their most intense point and how they end
  * We leveraged this by creating summary cards that highlight key takeaways, increasing the chances that users leave the app with a clear, memorable impression of the most important insights
:::

## {echarts4r} Implementation: Visual Result {background-color="#f9f9f9"}

::: {.r-stack style="text-align:center"}
![](img/echarts4r-ex1.png){width="75%"}

![](img/echarts4r-ex2.png){.fragment width="75%"}

![](img/echarts4r-ex3.png){.fragment width="75%"}

![](img/echarts4r-ex4.png){.fragment width="75%"}
:::

::: {.notes}
**Complete Dashboard**: Show integration of multiple principles

- For each image as it appears:
  * First image: "Gauge charts provide instant comprehension"
  * Second image: "The framing toggle changes perspective while maintaining data integrity"
  * Third image: ""Summary cards implement the Peak-End Rule"
- "Also, the card-based layout creates clear visual hierarchy and reduces cognitive load"
- "Every element in this dashboard has a specific psychological purpose"
- "This is what we mean by behavior-informed design - intentional choices based on how humans think"
:::

## {echarts4r} Implementation: Gauge Chart {background-color="#f9f9f9"}

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: "2-21|23-30"

# Packages: echarts4r, bslib
# Create intuitive gauge chart (Stage 2: Processing Fluency)
create_gauge_chart <- function(value, title) {
  e_charts() |>
    e_title(title) |>
    e_gauge(
      value,
      startAngle = 180,
      endAngle = 0,
      detail = list(formatter = "{value}%"),
      axisLine = list(
        lineStyle = list(
          color = list(
            c(value/100, "#1c6d7d"),  # Filled portion
            c(1, "#e9e9e9")           # Empty portion
          ),
          width = 30
        )
      )
    )
}

# Usage in dashboard (or use purrr::map)
layout_column_wrap(
  width = 1 / 4,
  create_gauge_chart(mean(music_data$`Metric 1`), "Metric 1"),
  create_gauge_chart(mean(music_data$`Metric 2`), "Metric 2"),
  create_gauge_chart(mean(music_data$`Metric 3`), "Metric 3"),
  create_gauge_chart(mean(music_data$`Metric 4`), "Metric 4")
)
```

::: {.notes}
**Gauge Charts**: Focus on visual processing benefits.

- "Gauge charts implement processing fluency perfectly"
- "We're customizing the gauge to eliminate unnecessary complexity"
- "The color gradient visually reinforces the value"
- "This visual simplicity reduces cognitive load while maintaining information richness"
- "The function encapsulates this pattern for consistent application"
:::

## {echarts4r} Implementation: Context {background-color="#f9f9f9"}

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: "2-21|23-29"

# Packages: echarts4r, bslib
# Create benchmark bar (Stage 4: Anchoring Effect)
create_benchmark_bar <- function(value, title, benchmark = 50) {
  e_charts() |>
    e_title(title) |>
    e_bar(
      value,
      legend = list(show = FALSE),
      showBackground = TRUE,
    ) |>
    e_flip_coords() |>
    e_labels(position = "inside") |>
    e_y_axis(show = FALSE) |>
    e_x_axis(show = FALSE, min = 0, max = 100) |>
    e_hide_grid_lines() |>
    e_mark_line(
      data = list(yAxis = benchmark),
      lineStyle = list(color = "#dd8500", type = "dashed", width = 2),
      label = list(show = TRUE, formatter = "Benchmark")
    )
}

# Usage in dashboard (or use purrr::map)
layout_columns(
  width = 1 / 3,
  card(create_benchmark_bar(mean(music_data$popularity), "Popularity")),
  card(create_benchmark_bar(mean(music_data$danceability), "Danceability")),
  card(create_benchmark_bar(mean(music_data$energy), "Energy"))
)
```

::: {.notes}
**Reference Lines**: Highlight the Anchoring Effect implementation.

- This directly addresses the Anchoring Effect from Stage 4"
- "Without reference lines, metrics like '65%' lack meaning - is it good or bad?"
- "Benchmarks provide the context users need for proper interpretation...and helps with alignment across users"
- "The distinctive orange dashed line creates visual separation from the data"
- "This simple addition dramatically improves decision quality"
:::

## {echarts4r} Implementation: Framing {background-color="#f9f9f9"}

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: "2-3|6-12|14-20|23-24|25-40|42-43|45-59"

# Packages: shiny, echarts4r, dplyr, purrr, bslib
# UI with framing toggle (Stage 4: Framing & Loss Aversion)
ui <- fluidPage(
  titlePanel("SoundPulse Dashboard"),

  # Stage 4: Framing toggle
  radioButtons(
    "framing", "Framing:", 
    choices = c("Progress" = "progress", "Gap" = "gap"),
    selected = "progress",
    inline = TRUE
  ),

  # Layout for charts using programmatically generated outputs
  layout_column_wrap(
    width = 1 / 3,
    card(echarts4rOutput("chart_popularity")),
    card(echarts4rOutput("chart_danceability")),
    card(echarts4rOutput("chart_energy"))
  )
)

# Server logic
server <- function(input, output, session) {
  # Reactive expression for framed data
  framed_data <- reactive({
    soundpulse_data |>
      mutate(
        display_value = if_else(
          input$framing == "progress",
          value,
          abs(benchmark - value)
        ),
        title = if_else(
          input$framing == "progress",
          metric,
          paste0(metric, " Gap")
        )
      )
  })

  # Grab KPI column names
  kpi_col_names <- c("popularity", "danceability", "energy")

  # Generate all chart outputs dynamically
  walk(kpi_col_names, function(kpi_col) {
    output_id <- paste0("chart_", kpi_col)

    output[[output_id]] <- renderEcharts4r({
      metric_data <- filter(framed_data(), metric == kpi_col)

      # Function from previous slide
      create_benchmark_bar(
        metric_data$display_value,
        metric_data$title,
        metric_data$benchmark
      )
    })
  })
}
```

::: {.notes}
**Framing Toggle**: Emphasize the powerful psychological effect.

- "This implementation shows how framing affects perception - same data, different decisions"
- "The 'Progress' frame emphasizes achievement - how far we've come"
- "The 'Gap' frame emphasizes what remains - how far we still need to go"
- "Different teams at SoundPulse used different frames - marketing used Progress, product used Gap"
- "The server-side code dynamically recalculates values based on the framing perspective"
- "This is psychological design directly impacting business decisions"
:::

## How Framing Affects Decision Making {background-color="#e6f0ff"}

:::: {.columns}

::: {.column .r-fit-text .fragment}
![](img/progress-frame.png){width="90%"}

::: {.nonincremental}
- **Progress (gain) Framing:**
  * Motivates by highlighting achievement
  * Creates positive momentum
  * Supports incremental improvement
:::
:::

::: {.column .r-fit-text .fragment}
![](img/gap-frame.png){width="90%"}

::: {.nonincremental}
- **Gap (loss) Framing:**
  * Creates urgency to address shortfall
  * Highlights areas needing improvement
  * May trigger risk-avoidance behaviors
:::
:::

::::

::: {.notes}
**Framing Effect**: Drive home the practical impact.

- This is a great example to illustrate the theory behind these concepts
- "In this case we are showing the *Same exact data*...but users experience completely different psychological impact"
- "Progress framing creates motivation and highlights achievement"
  - "Marketing preferred progress framing to build stakeholder confidence"
- "Gap framing creates urgency and focuses on what remains to be done"
  - "Product teams used gap framing to prioritize development"
- "This isn't just theory - a classic JDM study presented a 90% survival rate vs. a 10% mortality rate - people react completely differently"
- "This is why understanding psychological principles is so powerful for dashboard design"
- "Think about your own dashboards - are you framing data in a way that drives the right decisions?"
:::

## From Framework to Workflow: {bidux} {.center background-color="#e6f0ff"}

:::: {.columns .r-fit-text}

::: {.column width="56%" .nonincremental .fragment}
- **Current Features (Phase 1):**
  * 📚 Concept browser
    * `bid_concept("processing fluency")`
  * 🧠 BID stage functions
    * `bid_notice()` to `bid_validate()`
:::

::: {.column width="42%" .nonincremental .fragment}
- **Coming Soon (Phases 2-4):**
  * 🤖 LLM Integration 
  * 🛠️ UI Component Library
  * 📊 Testing and Validation Tools
:::

::::

::: {.notes}
**Package Introduction**: Generate excitement about practical tools.

- "We're not just sharing a framework or theories - we're giving you tools to implement it immediately"
- "The `{bidux}` package makes these principles accessible to every R developer"
- "Phase 1 is available today on GitHub"
- Future phases will include LLM integration and ready-to-use UI components"
- "Think of `{bidux}` as your behavioral science assistant for Shiny app development"
:::

## What {bidux} Looks Like Today {background-color="#f9f9f9"}

::: {.panel-tabset}

### Concepts

::: {.r-stack}
::: {.fragment .fade-out}
```{r}
#| echo: true
library(bidux)

# Concept dictionary with fuzzy searching and case-insensitive lookup
bid_concept("processing fluency")
```
:::

::: {.fragment .fade-in-then-out}
```{r}
#| echo: true

# Concept dictionary with fuzzy searching and case-insensitive lookup
bid_concept("processing fluency")$description
bid_concept("processing fluency")$implementation_tips
```
:::

::: {.fragment}
```{r}
#| echo: true

# Concept dictionary with fuzzy searching and case-insensitive lookup
viz_concepts <- bid_concepts("vis") 
paste("Found", nrow(viz_concepts), "visualization-related concepts")
```
:::
:::

### Notice

::: {.r-stack}
::: {.fragment .fade-out}
```{r}
#| echo: true

# BID Notice stage
notice_stage <- bid_notice(
  problem = "Users overwhelmed by filtering options and dropdown menus",
  # Theory parameter is optional - will auto-suggest appropriate theory if empty
  evidence = "Feedback shows 65% of users abandon the dashboard after first use"
)

notice_stage
notice_stage$theory
notice_stage$suggestions
```
:::

::: {.fragment}
```{r}
#| echo: true

# BID Notice stage
notice_stage <- bid_notice(
  problem = "Users overwhelmed by filtering options and dropdown menus",
  theory = "cognitive load",
  evidence = "Feedback shows 65% of users abandon the dashboard after first use"
)

notice_stage
notice_stage$theory
notice_stage$suggestions
```
:::
:::

### Chained Stages

::: {.r-stack}
::: {.fragment .fade-out}
```{r}
#| echo: true

# BID chained stages
project_docs <- notice_stage |>
  bid_interpret(
    central_question = "How can we simplify the filtering experience?",
    # More details are better, but the function will offer suggestions if needed
    data_story = list(
      hook = "Dashboard abandonment hurts decision-making",
      context = "Current UI requires too many decisions before seeing results",
      resolution = "Simplify filter UI with smart defaults",
      audience = "Marketing team",
      metrics = c("Time to insight", "Filter usage")
    )
  ) |>
  bid_structure(
    layout = "dual_process",
    concepts = c("Principle of Proximity", "Default Effect")
  )
```
:::

::: {.fragment}
```{r}
#| echo: true

# BID chained stages output
project_docs
project_docs$suggestions
```
:::
:::

### Components

::: {.r-stack}
::: {.fragment .fade-out}
```{r}
#| echo: true

# Get component suggestions based on your BID application
ui_suggestions <- bid_suggest_components(project_docs, package = "bslib")
ui_suggestions
```
:::

::: {.fragment .fade-in}
```{r}
#| echo: true

# Component suggestions output
ui_suggestions$component
ui_suggestions$description
ui_suggestions$code_example
```
:::
:::

:::

::: {.fragment}
More at: [github.com/jrwinget/bidux](https://github.com/jrwinget/bidux)
:::

::: {.notes}
**Code Example**: Show practical usage of the package

- "Let's look at what {bidux} offers today"
- "The concept dictionary provides implementation advice, not just theory"
- "You can search across psychology, UI/UX, and data visualization fields...and more!"
- "The workflow functions guide you through each BID stage"
- "Auto-suggestion of theories based on your problem description"
- "Component suggestions provide concrete UI implementation ideas"
- When the link appears: "We're launching this at ShinyConf today - check out the repo for full documentation"
:::

## Real-World Impact: SoundPulse Results {.center background-color="#4682b4"}

:::: {.columns}
::: {.column width="50%" .fragment}
> "The BID framework transformed how we approach our dashboards. What once took weeks of refinement now has clear direction from day one."
> 
> — Maya Chen, SoundPulse Research Director
:::

::: {.column width="50%" .nonincremental .fragment}
- Dramatically faster decision-making
- Substantial increase in insights per session
- Stakeholder satisfaction transformed from frustration to enthusiasm
- Significantly streamlined implementation of new metrics
:::
::::

::: {.notes}
**Results Summary**: Emphasize real-world impact.

- "The results speak for themselves"
- "Users who previously abandoned the dashboard now rely on it daily"
- "Decisions that took days now happen in minutes"
- We're currently building telemetry into our dashboards to gather more precise measurements...but the message from our users is already clear: this framework is changing how teams interact with their data
- When testimonial appears: "This transformation is what BID delivers"
- "Remember, this wasn't just about visual improvements - this was a fundamental shift in how our users...and developers...approached the dashboard"
:::

## Key Takeaways {.center background-color="#4682b4"}

::: {.nonincremental}
1. **BID** helps reduce friction and improve decision-making
2. **`{reactable}` + `{echarts4r}`** help bring BID to life
3. **`{bidux}`** supports you in applying BID at every stage<br>(try it today!)
4. Build dashboards that guide — not fry — your users
:::

::: {.notes}
**Wrap-Up**: Reinforce key messages with energy.

- "BID reduces friction between users and insights"
- "The tools we've shown today give you everything you need to get started"
- "`{bidux}` is available now to help structure your implementation"
- "Every dropdown you eliminate helps someone make a better decision"
- "This is about making your users successful, not just making dashboards"
:::

## {data-background-image="img/thank-you.png" data-background-size="100% 100%"}

<!-- TODO: replace background image with css and brand elements in img -->

<br>
<br>

:::: {.columns}
::: {.column width="55%" .r-fit-text}
### Milena Eickhoff {style="color: white;"}
![](img/headshot-eickhoff.jpg){width="80px" height="80px" style="border-radius: 50%;"}
![](img/qr-linkedin-eickhoff.svg){width="80px" height="80px"}

### Jeremy Winget, PhD {style="color: white;"}
![](img/headshot-winget.jpg){width="80px" height="80px" style="border-radius: 50%;"}
![](img/qr-linkedin-winget.svg){width="80px" height="80px"}

![](img/lets-connect.png){width="50%"}
:::
::::

::: {.notes}
**Q&A Setup**: Prepare for questions.

- We'd love to hear about your dashboard challenges"
- "Don't forget to follow us on LinkedIn/GitHub for updates on {bidux} development"
- "Feel free to connect with us for any questions or deeper discussions!"
- Quick answers to likely questions:
  * "How do you measure success?" → "Start with user testing and session analytics"
  * "Does this work for technical audiences?" → "Absolutely, technical users appreciate clarity too"
  * "How long does implementation take?" → "Start small - even one principle makes a difference"
:::