### 2025-04-08 conference-materials/2025_shiny-conf/slides.qmd
Updated notes for Milena's slides

---
title: "Death By Dropdown?"
subtitle: "A Developer's Guide To Building Dashboards That Won't Fry Your Client's Brain"
author: "Milena Eickhoff & Jeremy Winget, PhD"
format: 
  revealjs:
    title-slide-attributes:
      data-background-image: "img/title.png"
      data-background-size: "100% 100%"
    theme: simple
    slide-number: false
    footer: "ShinyConf 2025 | Death By Dropdown?"
    incremental: true
    transition: slide
    background-transition: fade
    highlight-style: github
    code-block-background: true
---

## The Problem {.center background-color="#4682b4"}

<!-- TODO: replace title background image with css and brand elements -->

:::: {.columns}
::: {.column .nonincremental width="60%"}
<h3>Ever felt like this?</h3>

- Endless dropdown menus
- Overwhelming complexity
- Users disengaged
- Delayed decision-making
- "Dashboard rot"
:::

::: {.column width="40%"}
![](https://media0.giphy.com/media/v1.Y2lkPTc5MGI3NjExbzVhM3d3MjJvYTNyc3puM2E0cmduZjFweXhrczNwcXI0b3FhN203aCZlcD12MV9pbnRlcm5hbF9naWZfYnlfaWQmY3Q9Zw/l44Q5OXJ6qaNr838Q/giphy.gif)
:::
::::

::: {.notes}
- Have you ever felt like this? You were asked to develop a dashboard with endless dropdown menus, overwhelming complecity, that you know would leave users disengaged and searching for answers? 
-	Send us an emoji if you have felt like this. How many of you have even abandoned a dashboard because it was too complex?
:::

## It Doesn't Have To Be This Way {.center}

<br>

::: {.r-fit-text}
By shifting the focus from <span style="color:#d55e00">**cramming data**</span> to <span style="color:#009e73">**crafting stories**</span>
:::

:::: {.columns .fragment}
::: {.column width="60%"}
<br>

SoundPulse's market research team was drowning in data, unable to extract meaningful insights from their complex dashboards.
:::

::: {.column width="40%"}
<br>

![](img/soundpulse-logo.png){width="100%"}
:::
::::

::: {.notes}
**Transition to Solution**: Shift to a hopeful, inspiring tone here.

- Don’t worry. It doesn’t have to be this way! We will give you a guide for crafting stories with data. The focus will not be on the data analysis itself but how you can update your ui to create dashboards that engage and empower the user.
- Let’s take SoundPulse for example, a composite client of us (a mix of different experiences we had with clients). It is a music streaming service, whos marketing team was faced with the exact same problem of too much data and too complex dashboards. 
- Let’s see how we can update their dashboards using the framework we developed and how you can implement our recommendations right away into your next project.
:::

## BID In Action: Dashboard Transformation {.center background-color="#f2f2f2"}

::: {.absolute top="80%" left="12%" width="80%"}
![](img/before-bid.png)
:::

::: {.notes}
**Show Impact**: Build frustration recognition with the "before" dashboard.

- Even though SoundPulse is a composite client...this is representative of many dashboards we've seen across the market research industry
- Point out key issues:
  * "Filter overload before seeing ANY insights"
  * "Multiple tabs with no clear starting point"
  * "Complete cognitive overload"
- "This is why dashboards get abandoned - nobody wants this experience"
- "Throughout the presentation, we'll show you the transformation process and the principles that guided our decisions"
:::

## Behavior Insight Design (BID) Framework {background-color="#e6f0ff"}

::: {.absolute left="7%" width="90%"}
![](img/bid-framework.png)
:::

:::: {.columns .absolute top="65%" left="4%" .r-fit-text}

::: {.column width="38%"}
1. <span style="color:#1c6d7d">**Notice**</span> the Problem
:::

::: {.column width="46%"}
2. <span style="color:#35a4ae">**Interpret**</span> the User's Need
3. <span style="color:#ff9000">**Structure**</span> the Dashboard
4. <span style="color:#ffca90">**Anticipate**</span> User Behavior
:::

::: {.column width="16%"}
5. <span style="color:#98d3ca">**Validate**</span> & Empower the User
:::

::::

::: {.notes}
**Framework Introduction**: Slow slightly to introduce - but drive home in next slide.

- "The 5 stages form an easy workflow: Notice → Interpret → Structure → Anticipate → Validate"
  * Click through
- "Notice is about identifying objective friction points - where users struggle with the interface"
- "Interpret goes deeper to understand what story or insight users truly need - moving from symptoms to diagnosis"
- "Structure focuses on organizing information in ways that align with how our brains process information"
- "Anticipate considers how cognitive biases might affect data interpretation"
- "Validate ensures users leave with clear takeaways and a positive final impression"
- "These stages are iterative - you'll often revisit earlier stages as you discover new insights"
- "This diagram will serve as our roadmap throughout the presentation"
:::

## Where BID Fits In {background-color="#f2f2f2"}

::: {.d-flex .align-items-center}
![](img/newton-portrait.jpg){width="15%" style="float:left; margin-right:15px;"}

> "If I have seen further, it is by standing on the shoulders of giants."
>
> — Isaac Newton
:::

:::: {.columns .absolute top="50%" .r-fit-text}

::: {.column .fragment }
- **UX Design**: Established best practices
- **Data Storytelling**: From data dumps to meaningful narratives
- **Psychological Science**: Evidence-based cognitive principles
- **Visual Communication**: Optimized for information perception
:::

::: {.column .callout-tip .fragment}
### Key Differentiator

Systematic integration of behavioral science within the natural dashboard development workflow
:::

::::

::: {.notes}
**Context Setting**: Position the framework within the broader fields.

- "BID integrates existing best practices from multiple disciplines"..."We are truly standing on the shoulders of giants here"
  * Click through example fields...UX, data storytelling, psych science, data/visual communication
- Key differentiator: "But what makes BID uniquely powerful is how it systematically connects psychological principles to specific coding decisions at each stage of development"
- "Unlike checklists that tell you what to do, BID guides how you think about your dashboard from psychological first principles"
- "The framework bridges theory and practice - showing not just what principles matter, but exactly when and how to apply them in your development workflow"
- "Even though there's a lot of theory, you don't need a psychology degree to apply these principles effectively"
- "To demonstrate...let's take a closer look at how we connected these stages to SoundPulse's dashboard."
:::

## SoundPulse Case Study: BID Stages 1-3 {background-color="#f9f9f9"}

:::: {.columns .nonincremental .r-fit-text}

::: {.column width="32%"}
- <span style="color:#1c6d7d">**Stage 1: Notice**</span>
  * Complex filtering overwhelmed users (*Cognitive Load*)
  * 72% users reported confusion (*Hick's Law*)
:::

::: {.column width="32%"}
- <span style="color:#35a4ae">**Stage 2: Interpret**</span>
  * Intuitive visualizations (*Processing Fluency*)
  * Clearer key insights<br>(*Data Storytelling*)
:::

::: {.column width="32%"}
- <span style="color:#ff9000">**Stage 3: Structure**</span>
  * Group related elements (*Principle of Proximity*)
  * Prioritization of metrics (*Dual-Processing Theory*)
:::

::::

::: {.absolute bottom="15%" left="15%" width="70%" height="30%"}
![](img/bid-framework.png)
:::

::: {.notes}
**Case Study Introduction:** Keep energy high while introducing concepts.

- **Stage 1 (Notice):** Users were overwhelmed with options
  * This is related to 'cognitive load' - which is the mental effort needed to process information
  * When the brain gets overtaxed, comprehension drops
  * Hick's Law also shows decision time increases logarithmically with the number of choices 
  * So more options, means exponentially more confusion
- **Stage 2 (Interpret):** Taking this to the next step helped us interpret what the user needed: Increased Processing Fluency
  * Processing fluency is the ease with which information is processed - when something is easy to process, people find it more credible and trustworthy
  * We also identified the key story the executives needed to understand...rather than showing all possible data
- **Stage 3 (Structure):** Moving to the Structuring stage...applied the 'Principle of Proximity' - which is a fancy way of saying we started grouping related elements together. 
  * Dual-Processing Theory distinguishes between fast, intuitive thinking and slower, deliberative thinking - we designed for both
  * We created groups of primary and secondary metrics to create a natural flow through the data
- "Now we'll show you the actual code implementation of these principles"
:::

## {reactable} Implementation: Visual Result {background-color="#f9f9f9"}

::: {.r-stack style="text-align:center"}
![](img/reactable-ex1.png){width="100%"}

![](img/reactable-ex2.png){.fragment width="100%"}

![](img/reactable-ex3.png){.fragment width="100%"}

![](img/reactable-ex4.png){.fragment width="100%"}

![](img/reactable-ex5.png){.fragment width="100%"}
:::

::: {.notes}
**Visual Impact**: Connect visuals to principles just mentioned.

- This is how we could transform the dashboard for SoundPulse using stages 1-3 of our framework. Here SoundPulse wanted us to display several metrics and KPIs for different songs and display them per subgroup – in this case age. 
  *	Notice how the grouped bars reduce cognitive load and the consistent legend simplifies comprehension. 
  *	Notice how color-coding instantly distinguishes the different subgroups. 
  *	Notice how the different shapes of metrics (bar charts) and KPIs (circles) instantly distinguishes the two groups. 
  *	And if a user wants to know more information drilldowns can be used for progressive disclosure. 
- This is the power of intentional design. Every element has a psychological purpose.

:::

## {reactable} Implementation: Basics {background-color="#f9f9f9"}

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: "4-5|6-18|19-23"

# Packages: reactable, dplyr, bslib
reactable(
  soundpulse_data,
  # Stage 1: Reduce cognitive load with grouped structure
  groupBy = "Song",
  # Stage 2: Enhance processing fluency with color coding
  columns = list(
    `Metric 1` = colDef(
      style = function(value) {
        # Color coding for instant comprehension
        case_when(
          value > 60 ~ list(background = "#1c6d7d", color = "white"),
          value > 40 ~ list(background = "#35a4ae", color = "white"),
          TRUE ~ list(background = "#ffca90", color = "black")
        )
      }
    )
  ),
  # Stage 1: Progressive disclosure reduces overload
  details = function(index) {
    song_details <- filter(details_data, song_id == soundpulse_data$Song[index])
    card(reactable(song_details, outlined = TRUE))
  }
)
```

::: {.notes}
**Basic Implementation**: Highlight the simplicity of applying BID.

- We promised you a guide on how you can implement our framework in your next project – so here it comes. 
- A package we can recommend for your next dashboard is reactable (also in combination with reactable formatter). 
- As a web developer I am scared of tables so I am glad to have a package that builds them for me and allows me to modify them easily. 
  * An option you can easily modify yourself is the groupBy  function – as we could see before grouping data variables on the dashboard reduces cognitive load while creating visual hierarchy that is easy to follow. 
  * Using the columns option (gives you a named list of column definitions where you can easily modify individual columns) you can easily color-code your table, for example depending on the value. It is not just aesthetic but creates instant processing fluency. 
  * You can use the details  option for progressive disclosure (the dynamic drilldowns on the previous slide). This code creates a bslib card component which encompasses a reactable that can be opened and closed. 
- This simple implementation already reduces cognitive load and enhances processing fluency but while those options are very easy to use we also want to go into more advanced options the reactable package gives you. 
:::

## {reactable} Implementation: Advanced {background-color="#f9f9f9"}

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: "2-10|12-16|17-27|29-38|39|40-56"

# Packages: reactable, purrr
# Define age group colors for consistency
age_colors <- c(
  "18-24" = "#1c6d7d",
  "25-34" = "#35a4ae", 
  "35-44" = "#98d3ca",
  "45-54" = "#ffca90",
  "55-64" = "#1c6d7d",
  "65-74" = "#dd8500"
)

# Create the table with dynamic legend component
div(
  # Main table component
  reactable(
    soundpulse_data,
    # Stage 3: Group related elements (Principle of Proximity)
    columnGroups = list(
      colGroup(
        name = "Core Metrics",
        columns = c("Metric 1", "Metric 2", "Metric 3", "Metric 4")
      ),
      colGroup(
        name = "Key Performance",
        columns = c("KPI 1", "KPI 2", "KPI 3")
      )
    )
  ),
  # Stage 1: Dynamic legend (Hick's Law - reduces choice complexity)
  div(
    style = css(
      display = "flex",
      align_items = "center",
      margin_top = "10px",
      padding = "8px",
      border_radius = "4px",
      background_color = "#f9f9f9"
    ),
    map(names(age_colors), function(age_group) {
      div(
        style = css(
          display = "flex", 
          align_items = "center", 
          margin_right = "15px"
        ),
        div(
          style = css(
            width = "10px", 
            height = "10px", 
            border_radius = "50%", 
            background_color = age_colors[age_group],
            margin_right = "5px"
          )
        ),
        span(age_group)
      )
    })
  )
)
```

::: {.notes}
**Advanced Features**: Highlight sophisticated techniques of package, but balance with BID principles.

- In the more advanced example we will show you how you can group columns and create a dynamic legend.
  * First, we create a vector that stores the colors we want to use to distinguish the subgroups. 
  * Then we create the reactable. You can use the columnGroups   option to group primary and secondary metrics together (example of the principle of proximity) – it creates a header above the columns. 
  * Below we create a container for the dynamic legend and map over the colors and age groups using a function from the purrr package. The dynamic legend reduces choice complexity for the user.
- Ask: “How many have used columnGroups in reactable before?” (engagement opportunity) 
  * If not many…great, we taught something!
  * If a lot…could also mention (quickly) how to extend with legend/div

:::

## SoundPulse Case Study: BID Stages 4-5 {background-color="#f9f9f9"}

:::: {.columns .nonincremental .r-fit-text}

::: {.column width="48%"}
- <span style="color:#ffca90">**Stage 4: Anticipate**</span>
  * Executives comparing metrics without context (*Anchoring Effect*)
  * Different perspectives needed for different teams (*Framing & Loss Aversion*)
:::

::: {.column width="48%"}
- <span style="color:#98d3ca">**Stage 5: Validate**</span>
  * Teams needed actionable summaries (*Peak-End Rule*)
:::
::::

::: {.absolute bottom="15%" left="15%" width="70%" height="30%"}
![](img/bid-framework.png)
:::

::: {.notes}
**Later stages:** Set up the psychological depth of later stages

- "For the later stages of the BID framework, we addressed more subtle psychological factors that affect how users interpret and act on data."
- **Stage 4 (Anticipate):** We noticed executives were making flawed comparisons without proper context.
  * This relates to the 'Anchoring Effect' - where people rely too heavily on the first piece of information they see
  * We also saw teams struggling to interpret the data in the same way, so we addressed this with something called a "framing toggle"...more on that in just a moment
- **Stage 5 (Validate):** Finally, we applied the 'Peak-End Rule' - the idea that people judge experiences mainly by their most intense moment and ending
  * We created summary cards highlighting key takeaways so users leave with a clear, memorable impression
:::

## {echarts4r} Implementation: Visual Result {background-color="#f9f9f9"}

::: {.r-stack style="text-align:center"}
![](img/echarts4r-ex1.png){width="75%"}

![](img/echarts4r-ex2.png){.fragment width="75%"}

![](img/echarts4r-ex3.png){.fragment width="75%"}

![](img/echarts4r-ex4.png){.fragment width="75%"}
:::

::: {.notes}
**Complete Dashboard**: Show integration of multiple principles

- For this dashboard SoundPulse wanted to visualize the statistics of different songs in more detail. Like the song Will it crash or run? By the Shiny Band. You might not have heard of that song yet but I’m sure you know what it’s about.  
  * We implemented benchmark lines to help with the anchoring effect, this value can easily be used by the user for comparison. 
  * We included a framing toggle button so that the SoundPulse marketing team can decide if they want to display their progress towards the benchmark or how much they are still missing to catch up with the benchmark. This button easily changes perspective while maintaining data integrity. 
  * Lastly, the summary card (that you could create yourself using a large language model) ensures that users leave with key insights. 
:::

## {echarts4r} Implementation: Gauge Chart {background-color="#f9f9f9"}

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: "2-21|23-30"

# Packages: echarts4r, bslib
# Create intuitive gauge chart (Stage 2: Processing Fluency)
create_gauge_chart <- function(value, title) {
  e_charts() |>
    e_title(title) |>
    e_gauge(
      value,
      startAngle = 180,
      endAngle = 0,
      detail = list(formatter = "{value}%"),
      axisLine = list(
        lineStyle = list(
          color = list(
            c(value/100, "#1c6d7d"),  # Filled portion
            c(1, "#e9e9e9")           # Empty portion
          ),
          width = 30
        )
      )
    )
}

# Usage in dashboard (or use purrr::map)
layout_column_wrap(
  width = 1 / 4,
  create_gauge_chart(mean(music_data$`Metric 1`), "Metric 1"),
  create_gauge_chart(mean(music_data$`Metric 2`), "Metric 2"),
  create_gauge_chart(mean(music_data$`Metric 3`), "Metric 3"),
  create_gauge_chart(mean(music_data$`Metric 4`), "Metric 4")
)
```

::: {.notes}
**Gauge Charts**: Focus on visual processing benefits.

- Another package we recommend for different graphs is the echarts4r package which provides a large selection of graphs that can be easily customized. 
- Gauge charts implement processing fluency perfectly and as you can see here you can modify the e_gauge function in many ways. 
  * We’re customizing the gauge to eliminate unnecessary complexity. 
  * The color gradient visually reinforces the value 
  * the detail function here gives you another option to further customize your chart. 
  * Layout_column_wrap by bslib can be used to display multiple graphs in a neat and adaptable way. 
:::

## {echarts4r} Implementation: Context {background-color="#f9f9f9"}

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: "2-21|23-29"

# Packages: echarts4r, bslib
# Create benchmark bar (Stage 4: Anchoring Effect)
create_benchmark_bar <- function(value, title, benchmark = 50) {
  e_charts() |>
    e_title(title) |>
    e_bar(
      value,
      legend = list(show = FALSE),
      showBackground = TRUE,
    ) |>
    e_flip_coords() |>
    e_labels(position = "inside") |>
    e_y_axis(show = FALSE) |>
    e_x_axis(show = FALSE, min = 0, max = 100) |>
    e_hide_grid_lines() |>
    e_mark_line(
      data = list(yAxis = benchmark),
      lineStyle = list(color = "#dd8500", type = "dashed", width = 2),
      label = list(show = TRUE, formatter = "Benchmark")
    )
}

# Usage in dashboard (or use purrr::map)
layout_columns(
  width = 1 / 3,
  card(create_benchmark_bar(mean(music_data$popularity), "Popularity")),
  card(create_benchmark_bar(mean(music_data$danceability), "Danceability")),
  card(create_benchmark_bar(mean(music_data$energy), "Energy"))
)
```

::: {.notes}
**Reference Lines**: Highlight the Anchoring Effect implementation.

- We implement the benchmark bar with an e_bar function that has only one value and add an e_mark_line. 
  * Here you can decide the width and color of the line and if it should be dashed or solid for example. The benchmarks provide the context users need for proper interpretation and creates visual separation from the data. This simple addition dramatically improves decision quality. 
  * Again, the layout columns from bslib can be used to align the graphs.
:::

## {echarts4r} Implementation: Framing {background-color="#f9f9f9"}

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: "2-3|6-12|14-20|23-24|25-40|42-43|45-59"

# Packages: shiny, echarts4r, dplyr, purrr, bslib
# UI with framing toggle (Stage 4: Framing & Loss Aversion)
ui <- fluidPage(
  titlePanel("SoundPulse Dashboard"),

  # Stage 4: Framing toggle
  radioButtons(
    "framing", "Framing:", 
    choices = c("Progress" = "progress", "Gap" = "gap"),
    selected = "progress",
    inline = TRUE
  ),

  # Layout for charts using programmatically generated outputs
  layout_column_wrap(
    width = 1 / 3,
    card(echarts4rOutput("chart_popularity")),
    card(echarts4rOutput("chart_danceability")),
    card(echarts4rOutput("chart_energy"))
  )
)

# Server logic
server <- function(input, output, session) {
  # Reactive expression for framed data
  framed_data <- reactive({
    soundpulse_data |>
      mutate(
        display_value = if_else(
          input$framing == "progress",
          value,
          abs(benchmark - value)
        ),
        title = if_else(
          input$framing == "progress",
          metric,
          paste0(metric, " Gap")
        )
      )
  })

  # Grab KPI column names
  kpi_col_names <- c("popularity", "danceability", "energy")

  # Generate all chart outputs dynamically
  walk(kpi_col_names, function(kpi_col) {
    output_id <- paste0("chart_", kpi_col)

    output[[output_id]] <- renderEcharts4r({
      metric_data <- filter(framed_data(), metric == kpi_col)

      # Function from previous slide
      create_benchmark_bar(
        metric_data$display_value,
        metric_data$title,
        metric_data$benchmark
      )
    })
  })
}
```

::: {.notes}
**Framing Toggle**: Emphasize the powerful psychological effect.

- For the framing toggle we use radio buttons that trigger the server-side code to dynamically recalculate the values based on the framing perspective. Different teams at SoundPulse used different frames - marketing used Progress, product used Gap. This is psychological design directly impacting business decisions . 
  * In the last part we loop over the different kpi columns (using a function from the purrr package) and update all charts dynamically with the function from the previous page.
:::

## How Framing Affects Decision Making {background-color="#e6f0ff"}

:::: {.columns}

::: {.column .r-fit-text .fragment}
![](img/progress-frame.png){width="90%"}

::: {.nonincremental}
- **Progress (gain) Framing:**
  * Motivates by highlighting achievement
  * Creates positive momentum
  * Supports incremental improvement
:::
:::

::: {.column .r-fit-text .fragment}
![](img/gap-frame.png){width="90%"}

::: {.nonincremental}
- **Gap (loss) Framing:**
  * Creates urgency to address shortfall
  * Highlights areas needing improvement
  * May trigger risk-avoidance behaviors
:::
:::

::::

::: {.notes}
**Framing Effect**: Drive home the practical impact.

- "This is a great example to illustrate the theory behind these concepts"
- "In this case we are showing the *Same exact data*...but users experience completely different psychological state"
- "Progress framing creates motivation and highlights achievement"
  - "Marketing preferred progress framing to build stakeholder confidence"
- "Gap framing creates urgency and focuses on what remains to be done"
  - "Product teams used gap framing to prioritize development"
- "This isn't just theory - a classic study in the Judgment and Decision Making literature presented identical medical treatment outcomes as either a 90% survival rate or a 10% mortality rate - people react completely differently to these framings"
- "This is why understanding psychological principles is so powerful for dashboard design"
- "Think about your own dashboards - are you framing data in a way that drives the right decisions?"
:::

## From Framework to Workflow: {bidux} {.center background-color="#e6f0ff"}

:::: {.columns .r-fit-text}

::: {.column width="56%" .nonincremental .fragment}
- **Current Features (Phase 1):**
  * 📚 Concept browser
    * `bid_concept("processing fluency")`
  * 🧠 BID stage functions
    * `bid_notice()` to `bid_validate()`
:::

::: {.column width="42%" .nonincremental .fragment}
- **Coming Soon (Phases 2-4):**
  * 🤖 LLM Integration 
  * 🛠️ UI Component Library
  * 📊 Testing and Validation Tools
:::

::::

::: {.notes}
**Package Introduction**: Generate excitement about practical tools.

- "We're not just sharing a framework or theories - we're giving you tools to implement it immediately"
- "The `{bidux}` package directly implements the 5-stage BID framework we've discussed today"
- "Each function maps to a specific stage and guides your thinking through that part of the development process"
- "Phase 1 is available today on GitHub - it provides the conceptual foundation and workflow guidance"
- "Future phases will include LLM integration to suggest specific implementations based on your needs"
- "We're also developing a UI component library with pre-built elements that embody these principles"
- "Think of `{bidux}` as your behavioral science assistant for Shiny app development"
:::

## What {bidux} Looks Like Today {background-color="#f9f9f9"}

::: {.panel-tabset}

### Concepts

::: {.r-stack}
::: {.fragment .fade-out}
```{r}
#| echo: true
library(bidux)

# Concept dictionary with fuzzy searching and case-insensitive lookup
bid_concept("processing fluency")
```
:::

::: {.fragment .fade-in-then-out}
```{r}
#| echo: true

# Concept dictionary with fuzzy searching and case-insensitive lookup
bid_concept("processing fluency")$description
bid_concept("processing fluency")$implementation_tips
```
:::

::: {.fragment}
```{r}
#| echo: true

# Concept dictionary with fuzzy searching and case-insensitive lookup
viz_concepts <- bid_concepts("vis") 
paste("Found", nrow(viz_concepts), "visualization-related concepts")
```
:::
:::

### Notice

::: {.r-stack}
::: {.fragment .fade-out}
```{r}
#| echo: true

# BID Notice stage
notice_stage <- bid_notice(
  problem = "Users overwhelmed by filtering options and dropdown menus",
  # Theory parameter is optional - will auto-suggest appropriate theory if empty
  evidence = "Feedback shows 65% of users abandon the dashboard after first use"
)

notice_stage
notice_stage$theory
notice_stage$suggestions
```
:::

::: {.fragment}
```{r}
#| echo: true

# BID Notice stage
notice_stage <- bid_notice(
  problem = "Users overwhelmed by filtering options and dropdown menus",
  theory = "cognitive load",
  evidence = "Feedback shows 65% of users abandon the dashboard after first use"
)

notice_stage
notice_stage$theory
notice_stage$suggestions
```
:::
:::

### Chained Stages

::: {.r-stack}
::: {.fragment .fade-out}
```{r}
#| echo: true

# BID chained stages
project_docs <- notice_stage |>
  bid_interpret(
    central_question = "How can we simplify the filtering experience?",
    # More details are better, but the function will offer suggestions if needed
    data_story = list(
      hook = "Dashboard abandonment hurts decision-making",
      context = "Current UI requires too many decisions before seeing results",
      resolution = "Simplify filter UI with smart defaults",
      audience = "Marketing team",
      metrics = c("Time to insight", "Filter usage")
    )
  ) |>
  bid_structure(
    layout = "dual_process",
    concepts = c("Principle of Proximity", "Default Effect")
  )
```
:::

::: {.fragment}
```{r}
#| echo: true

# BID chained stages output
project_docs
project_docs$suggestions
```
:::
:::

### Components

::: {.r-stack}
::: {.fragment .fade-out}
```{r}
#| echo: true

# Get component suggestions based on your BID application
ui_suggestions <- bid_suggest_components(project_docs, package = "bslib")
ui_suggestions
```
:::

::: {.fragment .fade-in}
```{r}
#| echo: true

# Component suggestions output
ui_suggestions$component
ui_suggestions$description
ui_suggestions$code_example
```
:::
:::

:::

::: {.fragment}
More at: [github.com/jrwinget/bidux](https://github.com/jrwinget/bidux)
:::

::: {.notes}
**Code Example**: Show practical usage of the package

- "Let's look at what {bidux} offers today"
- "The concept dictionary provides implementation advice, not just theory"
- "You can search across psychology, UI/UX, and data visualization fields...and more!"
- "The workflow functions guide you through each BID stage"
- "Auto-suggestion of theories based on your problem description"
- "Component suggestions provide concrete UI implementation ideas"
- "The package helps connect psychological principles to specific suggestions and components"
- "While we don't have time to go through a complete workflow demonstration, the package documentation on GitHub provides detailed examples"
- When the link appears: "We're launching this at ShinyConf today - check out the repo for full documentation"
:::

## Real-World Impact: SoundPulse Results {.center background-color="#4682b4"}

:::: {.columns}
::: {.column width="50%" .fragment}
> "The BID framework transformed how we approach our dashboards. What once took weeks of refinement now has clear direction from day one."
> 
> — Maya Chen, SoundPulse Research Director
:::

::: {.column width="50%" .nonincremental .fragment}
- Dramatically faster decision-making
- Substantial increase in insights per session
- Stakeholder satisfaction transformed from frustration to enthusiasm
- Significantly streamlined implementation of new metrics
:::
::::

::: {.notes}
**Results Summary**: Emphasize real-world impact.

- "The results speak for themselves"
- "While our quantitative metrics are still being collected through new telemetry systems we're implementing, qualitative feedback has been overwhelmingly positive"
- "Users who previously abandoned the dashboard now rely on it daily"
- "Decisions that took days now happen in minutes"
- It's still early, but the emerging message from our clients is clear: this framework is changing how teams interact with their data
- When testimonial appears: "This transformation is what BID delivers"
- "Remember, this wasn't just about visual improvements - this was a fundamental shift in how our users...and developers...approached the dashboard"
:::

## Key Takeaways {.center background-color="#4682b4"}

::: {.nonincremental}
1. **BID** helps reduce friction and improve decision-making
2. **`{reactable}` + `{echarts4r}`** help bring BID to life
3. **`{bidux}`** supports you in applying BID at every stage<br>(try it today!)
4. Build dashboards that guide — not fry — your users
:::

::: {.notes}
**Wrap-Up**: Reinforce key messages with energy.

- "BID reduces friction between users and insights by applying psychological principles in a systematic way"
- "The tools we've shown today - reactable and echarts4r - give you everything you need to get started"
- "The bidux package makes these principles accessible to every R developer, guiding you through each stage"
- "Start with small implementations of these principles - even one or two can dramatically improve your dashboard"
- "Every dropdown you eliminate or visual cue you add helps someone make a better decision"
- "This is about making your users successful, not just making dashboards"
:::

## {data-background-image="img/thank-you.png" data-background-size="100% 100%"}

<!-- TODO: replace background image with css and brand elements in img -->

<br>
<br>

:::: {.columns}
::: {.column width="55%" .r-fit-text}
### Milena Eickhoff {style="color: white;"}
![](img/headshot-eickhoff.jpg){width="80px" height="80px" style="border-radius: 50%;"}
![](img/qr-linkedin-eickhoff.svg){width="80px" height="80px"}

### Jeremy Winget, PhD {style="color: white;"}
![](img/headshot-winget.jpg){width="80px" height="80px" style="border-radius: 50%;"}
![](img/qr-linkedin-winget.svg){width="80px" height="80px"}

![](img/lets-connect.png){width="50%"}
:::
::::

::: {.notes}
**Q&A Setup**: Prepare for questions.

- We'd love to hear about your dashboard challenges"
- "Don't forget to follow us on LinkedIn/GitHub for updates on {bidux} development"
- "Feel free to connect with us for any questions or deeper discussions about implementation details we couldn't cover today"
- Quick answers to likely questions:
  * "How do you measure success?" → "Start with user testing and session analytics; we're building telemetry tools to help quantify impact"
  * "Does this work for technical audiences?" → "Absolutely, technical users appreciate clarity too - in fact, they often have the highest cognitive load due to the complexity of their work"
  * "How long does implementation take?" → "Start small - even one principle makes a difference. A complete BID implementation can take 1-2 sprints depending on dashboard complexity"
  * "How does this apply to domains beyond music?" → "We've successfully applied these principles in healthcare, finance, and retail dashboards - the psychological principles are universal"
:::